Mixed effect tweedie model in TMB

### 2019-04-25 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
tweaks to temp models

Load data
```{r}
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/data/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST.rds"))
nrow(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
unique(d_trawl$year)
d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth_m>40,0,1)
)
d_trawl[is.na(d_trawl$shallow), ]
d_trawl[is.na(d_trawl$SST_scaled), ]

glimpse(d_trawl)
```

Explore SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <60), aes(SST,temperature_c, colour = -depth_m)) + geom_point()
```

@@

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <50), aes(meanSST,temperature_c, colour = -depth_m)) + geom_point()
```

```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 130)
plot_spde(spde)
```

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + SST_scaled*shallow, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3b <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3
m_temp3_nosst
```

@@

nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth_m)

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```
Residual plots:
```{r}
ggplot(predictions3$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-7,8) +
  geom_smooth()
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-7,8) +
  geom_smooth()
```

>>>>

Load data
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# glimpse(all_sensor)
# 
# replace obviously faulty depth data with NA
all_sensor$depth_m[all_sensor$depth_m < 10] <- NA

all_depth <- all_sensor %>%
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

  # convert lat and lon to UTMs
  dplyr::mutate(X = longitude, Y = latitude) %>%
  gfplot:::ll2utm(., utm_zone = 9) %>%

  # interpolate missing depth values
  dplyr::rename(depth = depth_m) %>%
  gfplot:::interp_survey_bathymetry()

unique(all_depth$data$year)

d_trawl <- all_depth$data %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

hist(d_trawl$temperature_c)

# remove obvious sensor fail, d_trawl %>% filter(ssid==4, temperature_c >12), fishing event id = 481861
d_trawl <- d_trawl[d_trawl$fishing_event_id != 481861, ]
d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  #depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  #depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1)
)
# d_trawl[is.na(d_trawl$shallow), ]
# d_trawl[is.na(d_trawl$SST_scaled), ]

glimpse(d_trawl)

d_trawl <- d_trawl %>% dplyr::filter(year>2003)
```

Explore SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```
```{r}
ggplot(d_trawl, aes(DOY, SST, colour = -depth)) + geom_point() + facet_wrap(~year) 
```

```{r}
ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()

```

@@

```{r}
ggplot(dplyr::filter(d_trawl, depth <35), aes(meanSST,temperature_c, colour = -depth)) + geom_point()
```

EXLORE TEMPERATURE at DEPTH

```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "temperature_c")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
```

```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 650)
plot_spde(spde)
```

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + shallow*poly(DOY_scaled,2), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3b <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3
m_temp3b
```

@@

nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```
Residual plots:
```{r}
ggplot(predictions3$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
  xlim(2,15) +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
   xlim(2,15) +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
```
